### 2019-04-30 doc/bilevel/CommonAbstract.tex
Adding draft documentation for pao.bilevel
I haven't yet confirmed that this compiles.  I'm not expecting we'll
use all of this text here.

We describe new capabilities for modeling bilevel programs within
the Pyomo modeling software.  These capabilities include new modeling
components that represent subproblems, modeling transformations for
re-expressing models with bilevel structure in other forms, and
optimize bilevel programs with meta-solvers that apply transformations
and then perform optimization on the resulting model.  We illustrate
the breadth of Pyomo's modeling capabilities for bilevel programs,
and we describe how Pyomo's meta-solvers can perform local and
global optimization of bilevel programs.

%\keywords{Python \and algebraic modeling language \and bilevel program}