### 2021-04-13 doc/sand/abstract.tex
Adding SAND document

PAO is a Python\sphinxhyphen{}based package for Adversarial Optimization.  The goal of
this package is to provide a general modeling and analysis capability for
bilevel, trilevel and other multilevel optimization forms that express
adversarial dynamics.  PAO integrates two different modeling abstractions:
\begin{enumerate}
\sphinxsetlistlabels{\arabic}{enumi}{enumii}{}{.}%
\item {}
\sphinxstylestrong{Algebraic models} extend the modeling concepts in the
\sphinxhref{https://github.com/Pyomo/pyomo}{Pyomo} algebraic modeling language
to express problems with an intuitive algebraic syntax.  Thus, we
expect that this modeling abstraction will commonly be used by PAO
end\sphinxhyphen{}users.

\item {}
\sphinxstylestrong{Compact models} express objective and constraints in a manner
that is typically used to express the mathematical form of these
problems (e.g. using vector and matrix data types).  PAO defines
custom \sphinxstyleemphasis{Multilevel Problem Representations} (MPRs) that simplify the
implementation of solvers for bilevel, trilevel and other multilevel
optimization problems.

\end{enumerate}