### 2021-04-13 doc/sand/acknow.tex

We thank Richard Chen, John Siirola and Jean-Paul Watson for collaborating
on the development of \textbf{pyomo.bilevel}.  The PAO software is
derived from ideas we prototyped in \textbf{pyomo.bilevel}.  We also
thank members of the SECURE Grand Challenge LDRD who provided input on
design and application of PAO, especially Bryan Arguello, Jared Gearhart,
Ali Pinar, and Cindy Phillips .

This work was funded by the U.S. Department of Energy National Nuclear
Security Administration's Office of Defense Nuclear Nonproliferation
Research and Development (NA-22). Sandia National Laboratories is a
multimission laboratory managed and operated by National Technology
and Engineering Solutions of Sandia, LLC., a wholly owned subsidiary
of Honeywell International, Inc., for the U.S. Department of Energy's
National Nuclear Security Administration under contract DE-NA0003525.
This paper describes objective technical results and analysis.
Any subjective views or opinions that might be expressed in the paper
do not necessarily represent the views of the U.S. Department of Energy
or the United States Government.